Toward dynamic models of combat or lack thereof in animal mating

### 2015-12-22 ms.tex
Change growth model to exponential; density of maturing individuals
expoential much easier to solve! Little biological difference at this stage

\begin{document}
\mstitlepage
\noindent
% \doublespacing
\linenumbers

@@

The reproductive success of males in the population is determined via their ability to compete for and hold nests. Denote $N$ to be the total number of available nests, $R_i$ to be the reproductive value of nest $i$ (number of offspring per nest per year), $\bar{R}$ to be the average reproductive value of all nests (number of offspring per nest per year), and $p_{\rm n}(R)$ to be the probability density distribution of nest quality with respect to $R$. The distribution of nest qulatiies is then set according to the distribution
\begin{equation} \label{eq:pdf_R}
Pr(R_i = x) =\bar(R) \, p_{\rm n}(R).
\end{equation}

@@

Within the immature phase, we assume males grow with rate
\begin{equation} \label{eq:growth}
\frac{{\rm d} m}{{\rm d} t} = a m^{b}
\end{equation}
where $a$ and $b$ are metabolic constants. Following this equation, the size of individuals at time $t$, having started growing at time $0$ is then
\begin{equation} \label{eq:growth}
m(t) = \left(m(0)^{1-b} + t a(1-b)\right)^{\frac1{1-b}}.
\end{equation}

@@

\subsection{Numerical analysis}

\section{Figures}
\subsection {Figure 1. Event Algorithm}
Population state generates events, events are either:
\begin{enumerate}
    \item A new individual x has matured and entered the pool (based off of a trait(?))
    \item Discovery event wherin both a nest RR (repoductive resource) and an individual x are drawn from a pool, the metabolic costs of searching are deducted from the time differences between events, that nest is either:
\begin{enumerate}
    \item un-occupied: in which x takes ownership of RR.
    \item occupied by x' in which a constest occurs (see fig 2)
\end{enumerate}
\end{enumerate}
After the contest Energy costs are deducted from both the winner and the loser.
The winner also has a chance to abandon the RR (based off a trait).
If a death occurs (Energy(x) less than 0) then x is removed from the individual pool.

\subsection{Figure 2. Contest Algorithm}
for an occupier x' vs an invader x, there are 3 stages:
\begin{enumerate}
    \item Display 1
    \item Display 2
    \item Fight
\end{enumerate}
In the display phases the probability of escalation is dependant on the visible traits of x and x'
The rate of events increases as time t approaches t1

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/contest_algorithm}
\caption{Algorithm for modelling contests behaviour within the ``mature'' male population.}
\label{fig:events}
\end{figure}
\clearpage

>>>>

\begin{document}
\mstitlepage
\noindent
% \doublespacing
% \linenumbers

@@

The reproductive success of males in the population is determined via their ability to compete for and hold nests. Denote $N$ to be the total number of available nests, $R_i$ to be the reproductive value of nest $i$ (number of offspring per nest per year), $\bar{R}$ to be the average reproductive value of all nests (number of offspring per nest per year), and $p(R)$ to be the probability density distribution of nest quality with respect to $R$. The distribution of nest qulatiies is then set according to the distribution
\begin{equation} \label{eq:pdf_R}
Pr(R_i = x) =\bar{R} \, p(R).
\end{equation}

@@

Within the immature phase, we assume males grow exponentially with rate
\begin{equation} \label{eq:growth}
\frac{{\rm d} m}{{\rm d} t} = a \, m
\end{equation}
where $a$ is mass-based rate of growth. Integrating eq. \ref{eq:growth}, the size of individuals at time $t$, having started growing at time $0$ is given by
\begin{equation} \label{eq:growth}
m(t) = m(0)\exp(a \, t).
\end{equation}
If we assume a constant mortality rate $k_i$ during the immature phase, the probability of males surviving to time $t$ is given by
\begin{equation} \label{eq:surv_immature}
S(t) = \exp(-k_i \, t).
\end{equation}
The time $\tau$ taken for a male to grow from $m(0) \rightarrow m$ is then
\begin{equation} \label{eq:tau}
\tau(m_0, m) = \frac1{a}\ln\left(\frac{m}{m_0}\right).
\end{equation}
Assuming all remained in the immature phase, the density $n(m,t)$ of males with size $m$ at time $t$, would be given by \cite{DeRoos-1997}:
\begin{equation} \label{eq:n1}
n(m,t | {\rm immature}) = \frac{B(t - \tau(m_0, m))}{g(m)} \exp\left(-\int_{m_0}^m \frac{k_i}{g(m^\prime)} {\rm d}m^\prime \right),
\end{equation}
where $B(t)$ is the population birth rate at time $t$.
Substituting from above, we obtain
\begin{equation} \label{eq:n1}
n(m, t \, | {\rm immature}) = \frac{B(t - \tau(m_0, m))}{a\, m} \left(\frac{m}{m_0}\right)^{\frac{k_i}{a}}.
\end{equation}

Finally, we assume males transition from immature $\rightarrow$ mature according to logistic probability function
\begin{equation}\label{eq:allocation}
Pr({\rm mature} \, | m) = \frac1{1 + \exp\left(\alpha \left(1 - m / m_{\rm mat}\right)\right)},
\end{equation}
where $m_{\rm mat}$ is the size at which 50\% of the population is mature and $\alpha$ determines the width of the curve.

Thus we can calculate the flux of males size $m$ into the reproductive pool at time $t$ as
\begin{equation}\label{eq:flux}
n(m, t\, | {\rm immature}) \, \frac{{\rm d}}{{\rm d} m} Pr({\rm mature}\,  | m).
\end{equation}

@@

\subsection{Algorithms}

\subsection{Event Algorithm}
See Fig. \ref{fig:events}. Population state generates events, events are either:
\begin{enumerate}
    \item A new individual x has matured and entered the pool (based off of a trait(?))
    \item Discovery event wherin both a nest RR (repoductive resource) and an individual x are drawn from a pool, the metabolic costs of searching are deducted from the time differences between events, that nest is either:
\begin{enumerate}
    \item un-occupied: in which x takes ownership of RR.
    \item occupied by x' in which a constest occurs (see fig 2)
\end{enumerate}
\end{enumerate}
After the contest Energy costs are deducted from both the winner and the loser.
The winner also has a chance to abandon the RR (based off a trait).
If a death occurs (Energy(x) less than 0) then x is removed from the individual pool.

\subsection{Contest Algorithm}
See Fig. \ref{fig:contests}.
For an occupier x' vs an invader x, there are 3 stages:
\begin{enumerate}
    \item Display 1
    \item Display 2
    \item Fight
\end{enumerate}
In the display phases the probability of escalation is dependant on the visible traits of x and x'
The rate of events increases as time t approaches t1

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/contest_algorithm}
\caption{Algorithm for modelling contests behaviour within the ``mature'' male population.}
\label{fig:contests}
\end{figure}
\clearpage
